\blue{Numerical Investigation of Multi-Fluid Flows through Dual Saturated-Unsaturated Porous Media - change the title to something close to the main topic of the manuscript ... something with viscous flow instability or viscous fingering)}

### 2017-05-16 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updated
Updated (saved in .txt just in case)

% Copyright 2007, 2008, 2009 Elsevier Ltd
%
% This file is part of the 'Elsarticle Bundle'.
% ---------------------------------------------
%
% It may be distributed under the conditions of the LaTeX Project Public
% License, either version 1.2 of this license or (at your option) any
% later version.  The latest version of this license is in
%    http://www.latex-project.org/lppl.txt
% and version 1.2 or later is part of all distributions of LaTeX
% version 1999/12/01 or later c.
%
% The list of all files belonging to the 'Elsarticle Bundle' is
% given in the file `manifest.txt'.
%

@@

Due to complex and/or heterogeneous geological formations, detailed modelling and simulation of multiphase flows through porous media is crucial to understand the aforementioned processes. Description of physics and mechanisms of multiphase flows were reported by \citet{wooding_1976} with focus on capillary pressure and the flow regimes in capillary media. A comprehensive review of mechanisms of fingering (i.e., fluid instability) and forces acting on the interface between two immiscible fluids can be found in \citet{homsy_1987}. \citet{adler_1988} described both microscopic and macroscopic scale darcy-equations. Immiscible displacements in porous media with both capillary and viscous effects can be characterized by two dimensionless numbers, the 
capillary number $C$ (macroscopic level), which is the ratio of viscous forces to capillary forces, and the ratio $M$ (or MR for both microscopic and macroscopic level) of the two viscosities. For certain values of these numbers, either viscous or capillary forces dominate leading to the following basic concepts, (a) viscous fingering (this will be described in the following sections), (b) capillary fingering or (c) stable displacement (cross flow with piston shape front characteristics).
%In a porous medium that is statistically homogenous two immiscible fluids are asumed to coexist together within a solid matrix. In a microscopic level fluids are newtonian and incompressible and gravity is the only external force acting on the system. An interfacial tension $\sigma$ is assumed to exist at the interfaces separating the two fluids. The main difference between microscopic and macroscopic level is the concept of capillary pressure. Nevertheless we should refer to the area of flow instabilities  as these, first described by \citet{saffman_1958} %\red{(IS THIS TRUE? SAFFMANN MAY HAVE BEEN THE FIRST TO THROUGHLY DISCUSS VISCOUS INSTABILITIES, BUT THERE WERE OTHER SCIENTISTS THAT DESCRIBED FLUID INSTABILITIES BEFORE, E.G., RAYLEIGH(1896), TAYLOR (1955) ETC)}.
A distinction should also be made between microscopic and macroscopic instabilities. The first describe phenomena occurring locally, for example, Kelvin-Helmholtz and Saffman-Taylor instabilities \citep{saffman_1959}. On the other hand, macroscopic instabilities involve two-phase flow system \citep{bottoni_1992}. This paper will focus on microscopic instabilities and phenomena in two-phase systems while a this work will focus on a local flow instability due to viscous and stress forces, often referred as viscous instability or \textit{viscous fingering}.

>>>>

%  LocalWords:  checkboard
% Copyright 2007, 2008, 2009 Elsevier Ltd
%
% This file is part of the 'Elsarticle Bundle'.
% ---------------------------------------------
%
% It may be distributed under the conditions of the LaTeX Project Public
% License, either version 1.2 of this license or (at your option) any
% later version.  The latest version of this license is in
%    http://www.latex-project.org/lppl.txt
% and version 1.2 or later is part of all distributions of LaTeX
% version 1999/12/01 or later c.
%
% The list of all files belonging to the 'Elsarticle Bundle' is
% given in the file `manifest.txt'.
%

@@

Due to complex and/or heterogeneous geological formations, detailed modelling and simulation of multiphase flows through porous media is crucial to understand the aforementioned processes. Description of physics and mechanisms of multiphase flows were reported by \citet{wooding_1976} with focus on capillary pressure and the flow regimes in capillary media. A comprehensive review of mechanisms of fingering (i.e., fluid instability) and forces acting on the interface between two immiscible fluids can be found in \citet{homsy_1987}. \citet{adler_1988} described both microscopic and macroscopic scale darcy-equations. Immiscible displacements in porous media with both capillary and viscous effects can be characterized by two dimensionless numbers, the 
capillary number $C$ (macroscopic level), which is the ratio of viscous forces to capillary forces, and the ratio $M$ (or MR for both microscopic and macroscopic level) of the two viscosities. For certain values of these numbers, either viscous or capillary forces dominate leading to the following basic concepts, (a) viscous fingering (this will be described in the following sections), (b) capillary fingering or (c) stable displacement (cross flow with piston shape front characteristics).
%In a porous medium that is statistically homogenous two immiscible fluids are asumed to coexist together within a solid matrix. In a microscopic level fluids are newtonian and incompressible and gravity is the only external force acting on the system. An interfacial tension $\sigma$ is assumed to exist at the interfaces separating the two fluids. The main difference between microscopic and macroscopic level is the concept of capillary pressure. Nevertheless we should refer to the area of flow instabilities  as these, first described by \citet{saffman_1958} %\red{(IS THIS TRUE? SAFFMANN MAY HAVE BEEN THE FIRST TO THROUGHLY DISCUSS VISCOUS INSTABILITIES, BUT THERE WERE OTHER SCIENTISTS THAT DESCRIBED FLUID INSTABILITIES BEFORE, E.G., RAYLEIGH(1896), TAYLOR (1955) ETC)}.
 A distinction should be made between microscopic and macroscopic instabilities. Microscopic instabilities describe phenomena occurring locally, for example, Kelvin-Helmholtz and Saffman-Taylor instabilities \citep{saffman_1959}. On the other hand, macroscopic instabilities involve two-phase flow system \citep{bottoni_1992}. This paper will focus on microscopic instabilities and phenomena in two-phase systems while a this work will focus on a local flow instability due to viscous and stress forces, often referred as viscous instability or \textit{viscous fingering}.
